\begin{abstract}
We study the design of airline hub-and-spoke networks through a continuous optimization framework that avoids discrete network design variables while retaining the main economic and operational trade-offs of the problem. We first formulate a single-level multiobjective model that balances passenger-attraction and operating considerations. Passenger capture is modeled through entropy-based terms that induce a logit-type market share split between the proposed network and external travel alternatives, while operating decisions account for flight frequencies, link capacities, station capacities, and budget constraints. This lower-level formulation yields a smooth approximation of network design decisions and produces feasible hub-and-spoke topologies together with endogenous demand allocation.

To align the model with the airline's real economic target, we then develop a bilevel framework in which the upper level maximizes profit by tuning market-specific weights in the lower-level objective. The bilevel structure allows the lower-level continuous model to remain tractable while learning which origin-destination markets should receive more emphasis so that the induced design better matches profit-driven decisions. A central difficulty is the presence of fixed opening and hub costs, which introduce non-smooth behavior that is hard to embed directly in the bilevel problem. To address this issue, we promote sparsity in the capacity variables and use an iterative reweighted $\ell_1$ scheme to approximate fixed costs. At each iteration, the capacities obtained in the previous solve are used to update the reweighting coefficients, progressively sharpening the sparse pattern and improving the fidelity of the fixed-cost approximation.

The resulting methodology combines logit-based demand capture, continuous hub-and-spoke design, bilevel calibration of market weights, and reweighted sparsity for fixed-cost modeling within a unified optimization framework. This approach provides a practical alternative for large-scale airline network design settings in which exact mixed-integer bilevel formulations are difficult to solve, while preserving an explicit link between network structure, passenger choice, and profitability.
\end{abstract}
