\section{Bilevel Weight-Adjustment Framework}
The bilevel model treats the market-wise coefficients $\widehat{\alpha}^{od}$ and $\widehat{\beta}^{od}$ as upper-level decision variables. Their role is to determine how strongly each origin-destination market contributes to the lower-level passenger and entropy terms. The upper-level objective is the actual profit of the airline,
\begin{equation}
\Pi(x)=\sum_{(o,d)\in\mathcal{K}} p^{od}q^{od}f^{od}
-\sum_{(i,j)\in\mathcal{A}} c_{o_{ij}}a_{ij},
\end{equation}
while the lower-level response is the solution of the continuous network design problem introduced in the previous section.

Using the lower-level variable $y$ to emphasize the nested structure, the bilevel problem can be written as
\begin{align}
\max_{\widehat{\alpha},\widehat{\beta},x} \quad & \Pi(x) \\
\text{s.t.} \quad & x \in \arg\min_{y} \Phi_{\mathrm{LL}}(y;\widehat{\alpha},\widehat{\beta}), \\
& \underline{\alpha} \leq \widehat{\alpha}^{od} \leq \overline{\alpha},
\quad
\underline{\beta} \leq \widehat{\beta}^{od} \leq \overline{\beta}
\qquad \forall (o,d)\in\mathcal{K}.
\end{align}
The box constraints reflect the projected updates used in the implementation and keep the market hyperparameters in a numerically meaningful range.

\subsection*{Penalty reformulation}
Solving the bilevel problem directly is difficult because the lower-level solution set depends implicitly on the hyperparameters. Following the penalty-method viewpoint of recent bilevel optimization algorithms, we work instead with a penalized surrogate in which upper- and lower-level objectives are coupled in a single objective. For a penalty parameter $\gamma > 0$, we consider
\begin{equation}
\mathcal{P}_{\gamma}(x;\widehat{\alpha},\widehat{\beta})
=
-\Pi(x)
\;+\;
\gamma\,\Phi_{\mathrm{LL}}(x;\widehat{\alpha},\widehat{\beta})
\;+\;
\gamma\,\rho_{\mathrm{bud}}\,\mathrm{exceed}(x),
\end{equation}
where $\mathrm{exceed}(x)$ is the budget-violation slack and $\rho_{\mathrm{bud}}$ is a large positive coefficient. In the code, this is exactly the role played by the penalized objective solved in \texttt{cvx-ll.gms}: the profit term appears with the opposite sign, and the lower-level terms are added through a penalty coefficient denoted by \texttt{gamma}.

The interpretation is standard. For small values of $\gamma$, the penalized problem prioritizes the upper-level profit term. As $\gamma$ grows, the solution is pushed toward points that are increasingly consistent with lower-level optimality. This is the basic logic behind penalty methods for bilevel optimization, and it is the perspective that guides the algorithm used here.

\subsection*{Companion lower-level solve}
The implementation also solves the pure lower-level problem
\begin{equation}
\min_x \Phi_{\mathrm{LL}}(x;\widehat{\alpha},\widehat{\beta}),
\end{equation}
using \texttt{cvx-sl.gms}. This companion solve is needed because the market-hyperparameter updates are based on comparing the market contributions implied by the penalized bilevel step and by the lower-level step. In practice, the algorithm alternates between these two problems:
\begin{itemize}
\item a penalized bilevel step, solved through \texttt{cvx-ll.gms};
\item a lower-level step, solved through \texttt{cvx-sl.gms}.
\end{itemize}

\subsection*{Hyperparameter update}
After both solves are available, the market hyperparameters are updated by projected gradient-style steps. Denoting generically by $g_{\alpha}^{od}$ and $g_{\beta}^{od}$ the search directions obtained from the two solves, the update reads
\begin{align}
\widehat{\alpha}^{od}_{+} &= \Pi_{[\underline{\alpha},\overline{\alpha}]}
\left(\widehat{\alpha}^{od} - \mu_{\alpha} g_{\alpha}^{od}\right), \\
\widehat{\beta}^{od}_{+} &= \Pi_{[\underline{\beta},\overline{\beta}]}
\left(\widehat{\beta}^{od} - \mu_{\beta} g_{\beta}^{od}\right),
\end{align}
where $\Pi$ denotes Euclidean projection onto the admissible interval and $\mu_{\alpha},\mu_{\beta}$ are step sizes. In the current code, these projections are implemented by clipping \texttt{alfa\_od} and \texttt{beta\_od} after every update.

\subsection*{Resolution algorithm}
The complete method contains two nested loops. The outer loop updates the MM reweighting coefficients used to approximate fixed costs. The inner loop solves the penalized bilevel problem and updates the market hyperparameters. A compact description is given in Algorithm~\ref{alg:penalty-bilevel}.

\begin{algorithm}[t]
\caption{Penalty-based bilevel algorithm with outer sparsity updates}
\label{alg:penalty-bilevel}
\KwIn{Initial market weights $\widehat{\alpha}^{od},\widehat{\beta}^{od}$, initial reference capacities $s^{\mathrm{prev}},sh^{\mathrm{prev}},a^{\mathrm{prev}}$, step sizes $\mu_{\alpha},\mu_{\beta}$, penalty coefficient $\gamma$, numbers of outer and inner iterations.}
\For{$k=1,\dots,K_{\mathrm{outer}}$}{
  build the reweighted deployment term from $s^{\mathrm{prev}},sh^{\mathrm{prev}},a^{\mathrm{prev}}$\;
  \For{$t=1,\dots,K_{\mathrm{inner}}$}{
    solve the penalized bilevel problem $\min_x \mathcal{P}_{\gamma}(x;\widehat{\alpha},\widehat{\beta})$\;
    recover the penalized solution $x^{\mathrm{pen}}$\;
    update the logit-share bounds and related auxiliary quantities\;
    solve the pure lower-level problem $\min_x \Phi_{\mathrm{LL}}(x;\widehat{\alpha},\widehat{\beta})$\;
    recover the lower-level solution $x^{\mathrm{ll}}$\;
    compute the projected update directions for $\widehat{\alpha}$ and $\widehat{\beta}$ from $x^{\mathrm{pen}}$ and $x^{\mathrm{ll}}$\;
    update $\widehat{\alpha},\widehat{\beta}$ by projected gradient steps\;
  }
  set $(s^{\mathrm{prev}},sh^{\mathrm{prev}},a^{\mathrm{prev}})$ equal to the capacities of the latest penalized solution\;
}
\KwOut{Network capacities, hub structure, market shares, and learned market hyperparameters.}
\end{algorithm}

This algorithm should be understood as a practical penalty method rather than as an exact reformulation of the bilevel problem. Its purpose is to produce network designs that are simultaneously sparse, behaviorally meaningful, and aligned with the true profit objective of the airline.
