\section{Computational Scheme}
The computational procedure is best understood as the combination of two iterative ideas that already appear separately in the TFG and are coupled here in the airline implementation. The first idea is the majorization-minimization treatment of fixed costs through reweighted $\ell_1$ penalties. The second idea is the iterative adaptation of market coefficients in the bilevel layer.

For the deployment terms, the TFG replaces indicator-type fixed costs by reweighted linear penalties. In its generic form, a fixed activation term $\kappa\,\mathbf{1}_{\{z>0\}}$ is approximated at iteration $k$ through a contribution proportional to
\begin{equation}
\kappa \, \frac{z^{(k)}}{z^{(k-1)}+\varepsilon},
\end{equation}
which is the first-order MM surrogate associated with the logarithmic sparsity regularizer. This mechanism is the bridge between the discontinuous $\ell_0$ viewpoint and the continuous optimization model.

In the airline chapter of the TFG, this idea is used to represent deployment costs for airport, hub, and link capacities. In the current code, the reweighting is concentrated on the airport-capacity variables $s_i$ and $sh_i$, which are the dominant source of fixed-cost non-smoothness in the bilevel setting. The previous iterates \texttt{s\_prev}, \texttt{sh\_prev}, and \texttt{a\_prev} are written to GAMS before each solve and determine the linearized penalties used in the current iteration.

On top of that MM loop, the bilevel implementation in \texttt{generate\_data.py} introduces an inner calibration loop for the market-wise coefficients \texttt{alfa\_od} and \texttt{beta\_od}. The resulting algorithm alternates the following steps:
\begin{itemize}
\item solve the lower-level airline model with the current reweighting coefficients and current market weights;
\item recover capacities, market shares, and route flows from the GAMS output;
\item update auxiliary bounds related to feasible logit shares;
\item solve the companion single-level model used to evaluate the current design;
\item compute gradient-style corrections for \texttt{alfa\_od} and \texttt{beta\_od};
\item update the reference capacities used in the MM penalties and repeat.
\end{itemize}

This nested iteration is not an exact bilevel algorithm in the classical sense. Rather, it is a practical successive-approximation scheme that preserves the tractability of the continuous model while progressively improving two aspects of the solution: the fidelity with which fixed costs are represented, and the consistency between lower-level market emphasis and upper-level profit. That is precisely the computational philosophy inherited from the TFG and extended in the present project.
